---
title: "Activity Analysis"
author: "nana"
date: "6/2/2020"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv", header = TRUE)
activity$date <- as.Date(activity$date)
str(activity)
```

#### What is mean total number of steps taken per day?
The mean and median is
```{r}
total_every_day <- aggregate(activity["steps"], by=activity["date"], sum)
summary(total_every_day$steps)
```

Make a histogram of the total number of steps taken each day.
```{r}
hist(total_every_day$steps, main = "Histogram of Total Number of Steps per Day on  Impute Value",  xlab = "Total Number of Steps per Day", ylab = "Frequency", col = "purple", breaks = 30)
```

#### What is the average daily activity pattern?
Get mean of steps and plot time series
```{r}
time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(row.names(time_series), time_series, type = "l", xlab = "5-min interval", ylab = "Average across all Days", main = "Average number of steps taken", col = "pink", lwd=2)
```

Find which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
max_interval <- which.max(time_series)
names(max_interval)
```

#### Imputing missing values
The total number of missing values in the dataset
```{r}
sum(is.na(activity))
```
Replace with mean steps of 5 min interval
```{r}
activity$steps[is.na(activity$steps)] <- mean(activity$steps, na.rm = T)
sum(is.na(activity))
```
####New total number of steps 

```{r}
total_every_day <- aggregate(activity["steps"], by=activity["date"], sum)
summary(total_every_day$steps)
hist(total_every_day$steps, main = "Histogram of Total Number of Steps per Day on  Impute Value",  xlab = "Total Number of Steps per Day", ylab = "Frequency", col = "yellow", breaks = 30)
```

So after filling in NAs with mean, mean stay the same but the median is a bit different. And new data seems less skewed.

#### Are there differences in activity patterns between weekdays and weekends?
Label each day with "weekday" and "weekend", get mean for each 5 min interval and two categories of days.
```{r}
activity$weekdays <- weekdays(activity$date)
activity$weeks[(activity$weekdays == "Saturday" | activity$weekdays == "Sunday")] <- "weekend"
activity$weeks[!(activity$weekdays == "Saturday" | activity$weekdays == "Sunday")] <- "weekdays"
new_time_series <- aggregate(steps ~ interval + weeks, data = activity, mean)
```
Plot 
```{r}
library(lattice)
xyplot(steps ~ interval | weeks, new_time_series, type = "l", layout = c(1, 2),xlab = "Interval", ylab = "Number of steps", lwd=2)
```

Comparing to his activity on weekend, this person walks more in the morning but less in the afternoon on weekdays.
